---
title: "meta-analysis_code"
output: word_document
---

# to do
- overall effect of temp on fitness and dispersal
- tetranychus cinnabarinus_PT, Tetranychus evansi calls low temp "control" - changed it to 'low' but can't re-run Rmd because of issues with analysis - Ines is re-running this; emailed her to ask her to make sure it's "low"
- tetranychus turkestani, cinnabarinus, urticae - check sample size with researchers - emailed to ask

```{r libraries}
library(tidyverse)
library(mgsub)
library(esc)
library(metafor)
library(lme4)
```

```{r import files, echo = F}

# main effect of temperature on fitness
fitness_main_list <- list.files(path = "./outputs", pattern = "output_fitness_TempMain")

d_fitness_main <- NULL
for(file in fitness_main_list){
  name <- gsub(".txt", "", file)
  df <- read.table(paste("./outputs/", file, sep = ""), header = T)
  df <- rownames_to_column(df, var = "term")
  df$study <- str_sub(name, 25)
  assign(name, df)
  d_fitness_main <- bind_rows(d_fitness_main, df)
}

d_sampleN <- read.csv("./outputs/Sample_sizes.csv", header = T)

mods_main_list <- list.files(path = "./outputs", pattern = "moderators_")

d_mods <- NULL
for(file in mods_main_list){
  name <- gsub(".txt", "", file)
  df <- read.table(paste("./outputs/", file, sep = ""), header = T)
  df$study <- str_sub(name, 12)
  assign(name, df)
  d_mods <- bind_rows(d_mods, df)
}

d_mods$species <- gsub("\\_.*", "", d_mods$study)

```

# Effects of temperature on fitness

## Pairwise temperature treatment comparisons

We calculated Hedge's g for the pairwise temperature treatment comparisons of fitness for all studies using the 'esc' package in R. For fitness proxies that were normally distributed, we used the *esc_mean_se function* to convert treatment means and standard error to Hedge's g. For binary fitness proxies (i.e., survival) we used the *convert_or2d* function to convert log-odds ratios to Hedge's g.

```{r, esfitness, echo = F}

# pairwise temperature treatment differences in fitness


## list emmeans datasets
fitness_emmeans_list <- list.files(path = "./outputs", pattern = "output_fitness_emmeans")

## estimate Hedge's g
d_g_fitness_OH <- NULL # dataframe of effect sizes for fitness for opt vs high
d_g_fitness_OL <- NULL # dataframe of effect sizes for fitness for opt vs low
for(file in fitness_emmeans_list){
  name <- gsub(".txt", "", file)
  df <- read.table(paste("./outputs/", file, sep = ""), header = T) #read all emmeans files in list
  df$study <- str_sub(name, 24)
  assign(name, df)
  
  if("emmean" %in% names(df)){ #if there is an emmeans column (meaning the fitness proxy is normal) then estimate Hedge's g from means/se
    g_fitness_OH <- as.data.frame(esc_mean_se(grp1m = subset(df, Temp.treatment == "opt")$emmean[1], 
                                              grp1se = subset(df, Temp.treatment == "opt")$SE[1], 
                                              grp1n = subset(d_sampleN, study == df$study[1])$Fitness.Nopt, 
                                              grp2m = subset(df, Temp.treatment == "high")$emmean[1], 
                                              grp2se = subset(df, Temp.treatment == "high")$SE[1], 
                                              grp2n = subset(d_sampleN, study == df$study[1])$Fitness.Nhigh, 
                                              es.type = "g", 
                                              study = df$study[1]))
    g_fitness_OMH <- as.data.frame(esc_mean_se(grp1m = subset(df, Temp.treatment == "opt")$emmean[1], 
                                               grp1se = subset(df, Temp.treatment == "opt")$SE[1], 
                                               grp1n = subset(d_sampleN, study == df$study[1])$Fitness.Nopt, 
                                               grp2m = subset(df, Temp.treatment == "medium.high")$emmean[1], 
                                               grp2se = subset(df, Temp.treatment == "medium.high")$SE[1], 
                                               grp2n = subset(d_sampleN, study == df$study[1])$Fitness.Nmediumhigh, 
                                               es.type = "g", 
                                               study = df$study[1]))
    g_fitness_OL <- as.data.frame(esc_mean_se(grp1m = subset(df, Temp.treatment == "opt")$emmean[1], 
                                              grp1se = subset(df, Temp.treatment == "opt")$SE[1], 
                                              grp1n = subset(d_sampleN, study == df$study[1])$Fitness.Nopt, 
                                              grp2m = subset(df, Temp.treatment == "low")$emmean[1], 
                                              grp2se = subset(df, Temp.treatment == "low")$SE[1], 
                                              grp2n = subset(d_sampleN, study == df$study[1])$Fitness.Nlow, 
                                              es.type = "g", 
                                              study = df$study[1]))
    g_fitness_OML <- as.data.frame(esc_mean_se(grp1m = subset(df, Temp.treatment == "opt")$emmean[1], 
                                               grp1se = subset(df, Temp.treatment == "opt")$SE[1], 
                                               grp1n = subset(d_sampleN, study == df$study[1])$Fitness.Nopt, 
                                               grp2m = subset(df, Temp.treatment == "medium.low")$emmean[1], 
                                               grp2se = subset(df, Temp.treatment == "medium.low")$SE[1], 
                                               grp2n = subset(d_sampleN, study == df$study[1])$Fitness.Nmediumlow,
                                               es.type = "g", 
                                               study = df$study[1]))
    
    d_g_fitness_OH <- rbind(d_g_fitness_OH, cbind(g_fitness_OH, "Temp.contr" = "OH"), cbind(g_fitness_OMH, "Temp.contr" = "OMH"))
    d_g_fitness_OL <- rbind(d_g_fitness_OL, cbind(g_fitness_OL, "Temp.contr" = "OL"), cbind(g_fitness_OML, "Temp.contr" = "OML"))
  }
  
  if("response" %in% names(df)){ #if there is an response column (meaning the transformed fitness proxy is normal) then estimate Hedge's g from means/se
    g_fitness_OH <- as.data.frame(esc_mean_se(grp1m = subset(df, Temp.treatment == "opt")$response[1], 
                                              grp1se = subset(df, Temp.treatment == "opt")$SE[1], 
                                              grp1n = max(1, subset(d_sampleN, study == df$study[1])$Fitness.Nopt), 
                                              grp2m = subset(df, Temp.treatment == "high")$response[1], 
                                              grp2se = subset(df, Temp.treatment == "high")$SE[1], 
                                              grp2n = max(1, subset(d_sampleN, study == df$study[1])$Fitness.Nhigh), 
                                              es.type = "g", 
                                              study = df$study[1]))
    g_fitness_OMH <- as.data.frame(esc_mean_se(grp1m = subset(df, Temp.treatment == "opt")$response[1], 
                                               grp1se = subset(df, Temp.treatment == "opt")$SE[1], 
                                               grp1n = max(1, subset(d_sampleN, study == df$study[1])$Fitness.Nopt), 
                                               grp2m = subset(df, Temp.treatment == "medium.high")$response[1], 
                                               grp2se = subset(df, Temp.treatment == "medium.high")$SE[1], 
                                               grp2n = max(1, subset(d_sampleN, study == df$study[1])$Fitness.Nmedium.high), 
                                               es.type = "g", 
                                               study = df$study[1]))
    g_fitness_OL <- as.data.frame(esc_mean_se(grp1m = subset(df, Temp.treatment == "opt")$response[1], 
                                              grp1se = subset(df, Temp.treatment == "opt")$SE[1], 
                                              grp1n = max(1, subset(d_sampleN, study == df$study[1])$Fitness.Nopt), 
                                              grp2m = subset(df, Temp.treatment == "low")$response[1], 
                                              grp2se = subset(df, Temp.treatment == "low")$SE[1], 
                                              grp2n = max(1, subset(d_sampleN, study == df$study[1])$Fitness.Nlow), 
                                              es.type = "g", 
                                              study = df$study[1]))
    g_fitness_OML <- as.data.frame(esc_mean_se(grp1m = subset(df, Temp.treatment == "opt")$response[1], 
                                               grp1se = subset(df, Temp.treatment == "opt")$SE[1], 
                                               grp1n = max(1, subset(d_sampleN, study == df$study[1])$Fitness.Nopt), 
                                               grp2m = subset(df, Temp.treatment == "medium.low")$response[1], 
                                               grp2se = subset(df, Temp.treatment == "medium.low")$SE[1], 
                                               grp2n = max(1, subset(d_sampleN, study == df$study[1])$Fitness.Nmedium.low),
                                               es.type = "g", 
                                               study = df$study[1]))
    
    d_g_fitness_OH <- rbind(d_g_fitness_OH, cbind(g_fitness_OH, "Temp.contr" = "OH"), cbind(g_fitness_OMH, "Temp.contr" = "OMH"))
    d_g_fitness_OL <- rbind(d_g_fitness_OL, cbind(g_fitness_OL, "Temp.contr" = "OL"), cbind(g_fitness_OML, "Temp.contr" = "OML"))
  }
  
  if("prob" %in% names(df)){ #if there is a prob column (meaning the fitness proxy is binary) then estimate Hedge's g from OR
    d2 <- read.table(paste("./outputs/output_fitness_Summary_", df$study[1], ".txt", sep = ""), header = T)
    d2 <- rownames_to_column(d2, var = "term")
    
    g_fitness_OH <- as.data.frame(convert_or2d(or = exp(subset(d2, term == "Temp.treatmenthigh")$Estimate[1]), 
                                               se = subset(d2, term == "Temp.treatmenthigh")$Std..Error[1], 
                                               totaln = subset(d_sampleN, study == df$study[1])$Fitness.totalN, 
                                               es.type = "g", 
                                               study = df$study[1]))
    g_fitness_OMH <- as.data.frame(convert_or2d(or = exp(subset(d2, term == "Temp.treatmentmedium.high")$Estimate[1]), 
                                                se = subset(d2, term == "Temp.treatmentmedium.high")$Std..Error[1], 
                                                totaln = subset(d_sampleN, study == df$study[1])$Fitness.totalN, 
                                                es.type = "g", 
                                                study = df$study[1]))
    g_fitness_OL <- as.data.frame(convert_or2d(or = exp(subset(d2, term == "Temp.treatmentlow")$Estimate[1]), 
                                               se = subset(d2, term == "Temp.treatmentlow")$Std..Error[1], 
                                               totaln = subset(d_sampleN, study == df$study[1])$Fitness.totalN, 
                                               es.type = "g", 
                                               study = df$study[1]))
    g_fitness_OML <- as.data.frame(convert_or2d(or = exp(subset(d2, term == "Temp.treatmentmedium.low")$Estimate[1]), 
                                                se = subset(d2, term == "Temp.treatmentmedium.low")$Std..Error[1], 
                                                totaln = subset(d_sampleN, study == df$study[1])$Fitness.totalN, 
                                                es.type = "g", study = df$study[1]))
    
    g_fitness_OH$es <- g_fitness_OH$es*-1
    g_fitness_OMH$es <- g_fitness_OMH$es*-1
    g_fitness_OL$es <- g_fitness_OL$es*-1
    g_fitness_OML$es <- g_fitness_OML$es*-1
    d_g_fitness_OH <- rbind(d_g_fitness_OH, cbind(g_fitness_OH, "Temp.contr" = "OH"), cbind(g_fitness_OMH, "Temp.contr" = "OMH"))
    d_g_fitness_OL <- rbind(d_g_fitness_OL, cbind(g_fitness_OL, "Temp.contr" = "OL"), cbind(g_fitness_OML, "Temp.contr" = "OML"))
    

  }
}
```


```{r fitnessmeta}

# random effects metaanalytical model

m_fitnessOH <- rma(yi = d_g_fitness_OH$es, vi = d_g_fitness_OH$var)
m_fitnessOL <- rma(yi = d_g_fitness_OL$es, vi = d_g_fitness_OL$var)

```

```{r fitnessforest, echo = F, fig.width = 6.5}

forest(m_fitnessOH, slab = d_g_fitness_OH$study, addcred = TRUE, main = "Fitness: Opt vs Hot")

forest(m_fitnessOL, slab = d_g_fitness_OL$study, addcred = TRUE, main = "Fitness: Opt vs Cold")

```

```{r, esdispersaloverall, echo = F, fig.width = 6.5}

d_dispersal_overall <- read.csv("outputs/overall_dispersal_chisq_boot.csv")
m_dispersal.overall <- rma(yi = chisq, sei = se, data = subset(d_dispersal_overall, study != "Tetranychus.cinnabarinus_PT"))
forest(m_dispersal.overall, slab = subset(d_dispersal_overall, study != "Tetranychus.cinnabarinus_PT")$study, addcred = TRUE)


```

```{r, esdispersalcontr, echo = F}

# pairwise temperature treatment differences in dispersal

## list summary datasets
dispersal_summary_list <- list.files(path = "./outputs", pattern = "output_dispersal_Summary")

## estimate Hedge's g
d_OR_dispersal_OH <- NULL # dataframe of effect sizes for dispersal for opt vs high
d_OR_dispersal_OL <- NULL # dataframe of effect sizes for dispersal for opt vs low

for(file in dispersal_summary_list){
  name <- gsub(".txt", "", file)
  df <- read.table(paste("./outputs/", file, sep = ""), header = T) #read all summary files in list
  df <- rownames_to_column(df, var = "term")
  df$study <- str_sub(name, 26)
  assign(name, df)
  # extract OR
  OR_dispersal_OH <- data.frame("study" = df$study[1], 
                                "OR" = subset(df, term == "Temp.treatmenthigh")$Estimate[1], 
                                se = subset(df, term == "Temp.treatmenthigh")$Std..Error[1], 
                                "totaln" = subset(d_sampleN, study == df$study[1])$Dispersal.totalN)
  OR_dispersal_OMH <- data.frame("study" = df$study[1], 
                                 "OR" = subset(df, term == "Temp.treatmentmedium.high")$Estimate[1], 
                                 se = subset(df, term == "Temp.treatmentmedium.high")$Std..Error[1], 
                                 "totaln" = subset(d_sampleN, study == df$study[1])$Dispersal.totalN)
  OR_dispersal_OL <- data.frame("study" = df$study[1], 
                                "OR" = subset(df, term == "Temp.treatmentlow")$Estimate[1], 
                                se = subset(df, term == "Temp.treatmentlow")$Std..Error[1], 
                                "totaln" = subset(d_sampleN, study == df$study[1])$Dispersal.totalN)  
  OR_dispersal_OML <- data.frame("study" = df$study[1], 
                                 "OR" = subset(df, term == "Temp.treatmentmedium.low")$Estimate[1], 
                                 se = subset(df, term == "Temp.treatmentmedium.low")$Std..Error[1], 
                                 "totaln" = subset(d_sampleN, study == df$study[1])$Dispersal.totalN)
  
  d_OR_dispersal_OH <- rbind(d_OR_dispersal_OH, cbind("Temp.contr" = "OH", OR_dispersal_OH), cbind("Temp.contr" = "OMH", OR_dispersal_OMH))
  
  d_OR_dispersal_OL <- rbind(d_OR_dispersal_OL, cbind("Temp.contr" = "OL", OR_dispersal_OL), cbind("Temp.contr" = "OML", OR_dispersal_OML))

}

```


```{r dispersalcontrmeta, echo = T}

# outdated - using metaregression models below
# # join moderator dataset
# d_dispmeta_OH <- full_join(d_OR_dispersal_OH, d_mods)
# d_dispmeta_OL <- full_join(d_OR_dispersal_OL, d_mods)
# 
# # random effects metaanalytical model
# 
# m_dispersalOH <- rma(yi = OR, sei = se, mods = ~Taxonomic.group, data = subset(d_dispmeta_OH, study != "Tetranychus.turkestani" & study != "Tetranychus.cinnabarinus"))
# 
# d_dispmeta_OL$V_dispersal <- with(d_dispmeta_OL, I(se)^2)
# 
# m_dispersalOL <- rma.mv(yi = OR, V = V_dispersal, random = list(~ 1 | study), mods = ~Taxonomic.group, data = d_dispmeta_OL)

```



```{r fitdispcontr}

d_g_fitness <- rbind(d_g_fitness_OH, d_g_fitness_OL)
d_g_fitness$species <- gsub("\\_.*", "", d_g_fitness$study)
d_g_fitness <- d_g_fitness %>%
  rename(se_fitness = se)

d_OR_dispersal <- rbind(d_OR_dispersal_OH, d_OR_dispersal_OL)
d_OR_dispersal$species <- gsub("\\_.*", "", d_OR_dispersal$study)

d_OR_dispersal <- d_OR_dispersal %>%
  rename(se_dispersal = se)


d_fitdispcontr <- full_join(d_g_fitness, d_OR_dispersal, by = c("study", "Temp.contr"))
d_fitdispcontr$HILO <- with(d_fitdispcontr, ifelse(Temp.contr == "OH" | Temp.contr == "OMH", "hot", "cold"))
d_fitdispcontr <- full_join(d_fitdispcontr, d_mods)

# note -es so that when hot/cold have high fitness, "relative fitness" is greater
ggplot(subset(d_fitdispcontr, study != "Tetranychus.turkestani" & study != "Tetranychus.cinnabarinus"), aes(x = -es, y = OR, col = HILO, fill = HILO, size = 1/se_dispersal))+
  geom_point(pch = 1)+
  geom_smooth(method = "lm", alpha = 0.1)+
  geom_hline(yintercept = 0, lty = 2)+
  scale_y_continuous("Relative dispersal")+
  scale_x_continuous("Relative fitness")+
  scale_color_manual("", values = c("blue", "red"))+
  scale_fill_manual("", values = c("blue", "red"))+
  guides(size = "none")+
  theme_classic(base_size = 20)

# hot and cold together
# lmm

m_fitdispcontr <- lmer(OR ~ es+HILO + (1|Taxonomic.group/species) + (1|study), data = na.omit(subset(d_fitdispcontr, select = c(OR, es, HILO, Taxonomic.group, study, species), study != "Tetranychus.turkestani" & study != "Tetranychus.cinnabarinus")))
summary(m_fitdispcontr)
drop1(m_fitdispcontr, test = "Chisq")

# opt vs hot
# same study from different labs look like different study - need to correct
# lmm
m_fitdispcontr_OH <- lmer(OR ~ es + (1|Taxonomic.group/species) + (1|study), data = na.omit(subset(d_fitdispcontr, select = c(OR, es, Taxonomic.group, study, study, species), HILO == "hot" & study != "Tetranychus.turkestani" & study != "Tetranychus.cinnabarinus")))
summary(m_fitdispcontr_OH)
drop1(m_fitdispcontr_OH, test = "Chisq")

# meta regression
d_sub <- subset(d_fitdispcontr, HILO == "hot" & study != "Tetranychus.turkestani" & study != "Tetranychus.cinnabarinus")
d_sub$V_dispersal <- with(d_sub, I(se_dispersal)^2)

# plan for testing moderators:
##1) choose four moderators: 
### Testing biological hypotheses about what is driving dispersal: Temp variation, and Competition strength
### Testing hypotheses about methods that might influence estimate of effect size: Source, Dispersal.assay.duration
##2) remove if not significant
##3) with this dataset, only temperature variation is significant ***re-run with final dataset***
m_fitdispcontr_OH_meta <- rma.mv(yi = OR, V = V_dispersal, random = list(~ 1 | study, ~1|Taxonomic.group/species), mods = ~es + Temperature.variation, data = na.omit(subset(d_sub, select = c(OR, V_dispersal, study, Taxonomic.group, es, species, Temperature.variation))))

forest(m_fitdispcontr_OH_meta, slab = na.omit(subset(d_sub, select = c(OR, V_dispersal, study, Taxonomic.group, es, species, Temperature.variation)))$study, addcred = TRUE, main = "Dispersal: Opt vs Hot")
summary(m_fitdispcontr_OH_meta)
anova(m_fitdispcontr_OH_meta)


ggplot(subset(d_fitdispcontr, study != "Tetranychus.turkestani" & study != "Tetranychus.cinnabarinus" & HILO == "hot"), aes(x = es, y = OR, col = Temperature.variation, fill = Temperature.variation, size = totaln))+
  geom_point(pch = 1)+
  geom_smooth(method = "lm", alpha = 0.1)+
  geom_hline(yintercept = 0, lty = 2)+
  scale_y_continuous("Relative dispersal")+
  scale_x_continuous("Relative fitness")+
  scale_color_discrete("Temperature variation")+
  scale_fill_discrete("Temperature variation")+
  theme_classic(base_size = 20)

# opt vs cool

## lmm
m_fitdispcontr_OL <- lmer(OR ~ es + (1|Taxonomic.group/species) + (1|study), data = na.omit(subset(d_fitdispcontr, select = c(OR, es, Taxonomic.group, study, species), HILO == "cold" & study != "Tetranychus.turkestani" & study != "Tetranychus.cinnabarinus")))
summary(m_fitdispcontr_OL)
drop1(m_fitdispcontr_OL, test = "Chisq")

## meta regression
# plan for testing moderators:
##1) choose four moderators: 
### Testing biological hypotheses about what is driving dispersal: Temp variation, and Competition strength
### Testing hypotheses about methods that might influence estimate of effect size: Source, Dispersal.assay.duration
##2) remove if not significant
##3) with this dataset, only Source is significant ***re-run with final dataset***
m_fitdispcontr_OL_meta <- rma.mv(yi = OR, V = I(se_dispersal)^2, random = list(~ 1 | study, ~1|Taxonomic.group/species), mods = ~es + Source, data = na.omit(subset(d_fitdispcontr, HILO == "cold", select = c(OR, se_dispersal, study, Taxonomic.group, es, species, Source))))

summary(m_fitdispcontr_OL_meta)
anova(m_fitdispcontr_OL_meta)

ggplot(subset(d_fitdispcontr, study != "Tetranychus.turkestani" & study != "Tetranychus.cinnabarinus" & HILO == "cold" & es < 5 & es > 0), aes(x = es, y = OR, col = Source, fill = Source, size = totaln))+
  geom_point(data = subset(d_fitdispcontr, study != "Tetranychus.turkestani" & study != "Tetranychus.cinnabarinus" & HILO == "cold"), pch = 1)+
  geom_smooth(method = "lm", alpha = 0.1)+
  geom_hline(yintercept = 0, lty = 2)+
  scale_y_continuous("Relative dispersal")+
  scale_x_continuous("Relative fitness")+
  scale_color_discrete("Source")+
  scale_fill_discrete("Source")+
  theme_classic(base_size = 20)

forest(m_fitdispcontr_OL_meta, slab = na.omit(subset(d_fitdispcontr, HILO == "cold", select = c(OR, se_dispersal, study, Taxonomic.group, es, species)))$study, addcred = TRUE, main = "Dispersal: Opt vs Cool")


```

```{r}

```

