---
title: "Temperature dependent dispersal: A distributed experiment"
author: "Greta Bocedi, Caterina Berardo, Nicky Lustenhouwer, Lesley Lancaster"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r instructions, echo = F, warning = F}

# INSTRUCTIONS FOR RESEARCHERS

# Here, you'll find instructions and information to fill out about your species. Once you have finished following the instructions, go to the render_script in your species folder and run the code - this will automatically knit a word document to the shared Knitted_Markdowns folder. You don't need to make any changes to the analysis outside of this code chunk, but feel free to do so if it's appropriate for your species. If you make substantive changes to either of the models, please flag it for Celina.

# Note: Do not try to "run all" within R, you will get an error. Knitting the document from the RStudio menu will only save it to your species folder, not add it to the shared folder, so please use render_script.

# 0. Add names of researchers who worked on this species to author list above

# 1. Before starting, make sure your data files are named:
# Phase1_fitness_Genus.species.csv
# Phase2_dispersal_Genus.species.csv

## code below assumes .csv format (only in the code chunk named 'loaddata'; You can change this if you would rather not use .csv)


# 2. Install any missing libraries from the list below:
library(tidyverse)
library(lme4)
library(ggplot2)
library(emmeans)
library(DHARMa)

# 3. Fill out information about your species below
Genus = "Callosobruchus"
species = "maculatus"
fitness.proxy = "Fecundity"
low.temp = 22 #value of low temperature treatment in Celsius
opt.temp = 33 #value of optimal temperature treatment in Celsius
high.temp = 37 #value of high temperature treatment in Celsius


# 4. Check that the data files have the following headers (order of columns is not important):

## phase1 : Temp.treatment (levels = low, opt, high), Temp.value (in degrees C), random effects (name as appropriate), Fitness
### It's important to give the column of the fitness variable the header "Fitness". You have already provided a description of the fitness proxy which will be used in the knitted document.

## phase 2:

### If your phase 2 data are formatted with mesocosms as rows and number of dispersers/number of non-dispersers as columns, then these headers:
#### Temp.treatment (levels = low, opt, high), Temp.value (in degrees C), random effects (name as appropriate), Num.disp, Num.nondisp (or alternatively, Num.disp and total N)

### If your phase 2 data are formatted with individuals as rows and a dispersed column (0s and 1s) giving dispersal status for each individual, then these headers:
#### Temp.treatment (levels = low, opt, high), Temp.value (in degrees C), random effects (name as appropriate), Dispersed (0/1)

# note that the datasets can also include fixed effects and phenotypes, as appropriate for your study


# 5. Type of fitness data

## Answer questions below about the fitness proxy

### Is your fitness variable binary? (choose one)
fitness.binary = FALSE

### Is the fitness variable continuous normal (without transformation)? (choose one)
fitness.normal = TRUE

### Is the fitness variable normal after transformation? (choose one; choose NA if you said "TRUE" to either question above)
fitness.transformed.normal = FALSE

## If fitness variable is continuous normal AFTER transformation, choose transformation used (just write it in if not on this list):

transform.description <- "square-root/log/log10" # leave as-is if not applicable
# fitness.transformation.function <- function(x) {return (XYZ(x))} # give transformation formula here; replace XYZ with transformation e.g., sqrt(x) or log(x + 1), etc., where x will be Fitness, below

fitness.transformation.function <- function(x) {return (log(x))} # give transformation formula here; replace "log" with transformation used e.g., sqrt(x) or log(x + 1), etc., where x will be Fitness, below


## note: if fitness (or transformed fitness) does not have a binomial or normal distribution, check in with Celina.


# 6. Structure of dispersal data

## Is there overdispersion in the dispersal data? If you do not yet know, choose FALSE and this will be dealt with in step 12.

overdispersion = FALSE


# 7. Dealing with mortality in dispersal assay

## note on mortality: everyone will deal with mortality during the dispersal assay in a slightly different way. If individuals are rows in your dataset, you should have a row that indicates "Survived" (Y/N) through the entire experiment. In this case, set filter.dead = TRUE below. If mesocosms are rows in your dataset, you may have a column indicating the number of individuals that died. In this case, dead individuals can be excluded directly in the dispersal formula (e.g., (Num.disp, TotalN-(Num.disp+Num.dead))).

## Make sure to describe how you dealt with mortality in the methods below.

## If you are unsure how to handle mortality in your dataset, check with Celina.

## Do rows represent individuals in the dataset and are there dead individuals that should be removed from the dispersal data?

filter.dead = FALSE # choose FALSE if this does not apply to your species

# 8. Model formulae

## Formula for fitness: (fill in as appropriate)

fitness.formula <- "Fitness ~ Temp.treatment"

### notes:
#### TRANSFORM FITNESS VARIABLE (if needed) *WITHIN* THE FORMULA; doing it within the formula is important for contrasts below  
#### other.fixed.effects does not apply to every species; include as appropriate; fixed effects should be standardized
#### include as many random effects as appropriate
#### Fitness can be normal or binary, including 0s and 1s or cbind(number.survived, number.dead)

### If fitness is binary: (ignore if fitness is not binary)

### What do Fitness = 0 and Fitness = 1 mean?

Fit0 <- "Died/Survived" # or fill in other
Fit1 <- "Died/Survived" # or fill in other

## Formula for dispersal: (fill in as appropriate)
  
dispersal.formula <- "Dispersed ~ Temp.treatment + Length + (1|Group)"

### notes
#### other.fixed.effects does not apply to every species; include as appropriate; fixed effects should be standardized
#### include as many random effects as appropriate
#### Num.disp/Num.nondisp can be replaced with column "Dispersed" with values 0 and 1, if individuals are rows in your dataset
#### remember to deal with mortality directly in dispersl formula if appropriate


# 9. Moderators

## We are collecting data on all proposed moderators; we will use a subset of these in the analysis
## choose option on right hand side
## see description of moderators and levels in README.md in project repo

taxonomic.group <- "arthropod"
rearing.environment <- "treatmentT" ## GB: we have a fully factorial design  so it could be both commonT and treatment T
comp.strength <- "low/moderate/high" ## GB: need to check with Lesley
source <- "lab" # lab population or field caught
habitat <- "terrestrial"
Tvariation <- "homogenous"
species.body.size <- 3 # mean body length in mm
passive.dispersal <- "No"
inbred <- "mixed"
disp.assay.duration <- 0.001 # GB: considering a generation time of 3 weeks
# order of magnitude for dispersal assay duration / generation time, i.e., approximate number of generations the dispersal assay lasted
fitness.proxy.type <- "other"

# 10. Fill out methods information in text below.

```

## Author contributions
Possible categories:
- conceptualization 
- statistical analysis  
- methodology (design of methodology)   
- investigation/conduct experiment  
- funding acquisition  
- writing (original draft, supplement)
- list others as needed  

## Acknowledgements
Research assistants, funding sources, etc.

# Supplementary methods for *`r (print(paste(Genus, species)))`*

## Study organism
<!-- Name and strain/origin of species.  

Provide basic life history information (aquatic/terrestrial, mode of dispersal, lifespan, etc.).   

Briefly describe what is known about the temperature range the species experiences (geographic range, seasonality, type of habitat, etc.) in its natural system or the history of the conditions it has experienced in the lab.  

How were organisms sourced?  
-	Field caught: provide information on timing, geographic coordinates, type of habitat.  
-	Lab sourced: provide information on source.  
-	Provide license/ authorization numbers for catching animals in the field and/or doing experiments, where applicable. 

How were individuals/populations treated before the experiment?  
-	Rearing/holding conditions (temperature, food, light).  
-	How long were experimental subjects held before the experiment?  

Were specific individuals/phenotypes/stages selected for the experiment? --> 

Seed beetles (*Callosobruchus maculatus*), also known as the southern cowpea weevil.

*C. maculatus* is a seed beetle originated from Africa but is now spread across the tropics and sub-tropics (Kébé et al. 2017). Its ancestral host is *Vigna unguiculata* (black-eyed pea) but it is considered a major pest of many plants from the Fabaceae family such as adzuki or mung beans (Tuda et al. 2006). With agriculture, it became a stored-product pest as females lay their eggs on dry bean surface. After hatching, the larva will develop entirely inside the bean, digging a chamber causing severe damage to the bean. Adult beetles will emerge from the beans about 3-4 weeks after ovipositon at 27°C. Their life expectancy as adults at 27°C is about 8 to 12 days when not supplied with food or water.  

We used individuals from a labporatory stock population we created in April 2022 by mixing individuals from 2 populations sourced from agricultural fields in 2010 in Nigeria and Togo respectively (Berger et al. 2014). These 2 populations were maintained separately for approximatively a decade at 29°C in the laboratory. We kept the mixed population at 27°C, under a 12:12 (L:D) photoperiod, in controlled-temperature incubators. Beetles were reared on their ancestral host *V. unguiculata* (cowpea).



## Study design

### Temperature treatments

<!-- State the temperatures used for Topt, Thot, Tcold, and explain how they were chosen.  

State whether there was daily variation in temperature, and if so, describe (e.g., mean and standard deviation of temperature for each treatment). -->

Topt = 33 ± 0.5°C 
Thot = 37 ± 0.5°C 
Tcold = 22 ± 0.5°C

We chose these temperature based on previous assays performed by the group and literature. We aslo tested an additional temperature Tlab = 27 ± 0.5°C, as this is the temperature we keep the stock population at.

<!-- @Lelsley: do you have a more detailed explanation for the choice of temperature? -->


### Phase 1: Fitness assay

<!-- Describe the fitness proxy (growth rate, survival, body condition, reproductive output, etc.). If proxy is indirect (e.g., body condition), explain what is known about the relationship between the proxy and actual fitness in the species.  

 Provide details about the methods for estimating the fitness proxy including:  
-	Physical details about the experimental setup including size/shape/volume of experimental arenas, substrate/medium of experimental arenas.  
-	Replicates  
    - Specify number of experimental replicates  
    - Number of strains and number of replicates per strain  
    - Indicate if experimental blocks were used and the blocking design (number of blocks, how many replicates of each treatment per block, time/space between blocks)
-	Type and amount of food provided during experiment  
- Starting density  
-	Duration of fitness assay  
-	Controlled variables: food (ad lib?), water (ad lib?), sex ratio  
- Phenotypes measured (activity, body condition, body size, etc.) and how they were estimated --> 

A virgin male and a virgin female were collected from the stock population at 27°C, and placed in a petri dish for 24 hours to ensure mating. Different mating pairs were collected at the same time. After 24 hours, each female was located in a petri dish with 10g of beans to allow oviposition. After other 24 hours, the female was removed and the beans were located in the incubators at the 4 different temperatures (still in the petri-dishes). 20 replicate females per temperature were assayed. To ensure beetles were virgin at collection, after 10 days, we isolated beans in 48-well culture plates and we counted the eggs. Beans were checked daily as the emergence time approached and emerging offspring were counted daily (F1 hatching success).Only individuals found alone at emergence, or with other individuals of the same sex within a well, were used in the assay. A virgin F1 female from each replicate, younger than 24 h, was mated with a virgin male (taken from another plate at the same temperature) and the cycle was repeated to evaluate the fecundity of the generation reared at the treatment temperature (F2 fecundity). Seed beetles do not require to eat for mating so they were not supplied with food or water, as in many experimental set-ups (Miller and Inouye 2013; Downey et al. 2015; Ochocki et al. 2020).

We measured three fitness proxies: <!-- add rationale for it -->

1. F1 hatching success: number of offspring emerged / number of eggs laid at different temperature by parents reared and mated at 27 °C.

2. F2 fecundity: number of eggs laid by females reared at the treatment temperature. Fathers were also reared at the treatment temperature.

3. F2 hatching success: number of offspring emerged / number of eggs laid at different temperature by parents reared at the treatment temperature.

In this analyses we only consider the F2 fecundity as fitness proxy.

### Phase 2: Dispersal assay

<!-- Provide details about the methods for estimating dispersal including:  
-	Physical details about the experimental setup including size/shape/volume of experimental arenas, substrate/medium of experimental arenas, type of system (1 or 2 patches).  
-	Replicates  
    - Specify number of experimental replicates  
    - Number of strains and number of replicates per strain  
    - Indicate if experimental blocks were used and the blocking design (number of blocks, how many replicates of each treatment per block, time/space between blocks)  
-	Type and amount of food provided during experiment  
- Starting density  
-	Duration of dispersal assay including:  
    - time between birth and start of acclimation period  
    - duration of acclimation period  
    - timing and duration of dispersal assay  
-	Controlled variables: food (ad lib?), water (ad lib? if applicable), sex ratio  
- Phenotypes measured (activity, body condition, body size, etc.) and how they were estimated  
- Whether individuals used in phase 2 were the same as used in phase 1 
- Provide details on data collection including:  
    -	The sampling regime (time intervals between sampling).  
    -	Sampling method including whether individuals were marked and whether sampling was observed or automated. Please include validation if automated.  
- State whether mortality during the dispersal assay was recorded. Describe how mortality was handled for the statistical analysis (e.g., individuals that died during the dispersal assay were excluded from the analysis of dispersal).   -->

We applied a fully factorial design, collecting individuals reared at 4 different temperatures and assaying their dispersal propensity at each temperature. 

We assessed dispersal propensity and dispersal distance using linear mazes adapted from Ochocki and Miller (2017) experimental design. Mazes consisted of plastic tubs (volume: 75ml) connected by 20cm long plastic tubes (diameter: 0.5cm). Each tub contained 9g of cowpeas (~45 beans). All individuals (coming from 85 plates per temperature) were collected as virgin and aged not more than 24 hours old when assayed. 10 individuals (sex ratio 1:1) were introduced in the first tub. Each individual was sourced from different families to avoid pseudoreplication. We allowed an acclimation phase without dispersal of approximately 40 min (a quarter of the subsequent dispersal phase duration). After this amount of time, we opened the tube and allowed individuals to disperse for the 2 hours. After these 2 hours, we recorded the length (photographing them from 25 cm of distance and using ImageJ’ “Measure” function 3 times per individual), the sex and the position of each individual in the maze (Emigrated? Yes or No). If the individuals, after the 2h, were in between tubes, we have considered them as dispersed.
Assessments were made on 8 replicates (8 runs in linear maze) per each treatment, for a total of 128 replicates. The experiment was divided in 2 consecutive days (4 replicates per day). The individuals were not the same as the ones used for the fitness assay. Mortality did not occur during these assays.

```{r instructions2, echo = F}
# 11. Knit RMarkdown

## Now it is time to knit the RMarkdown document. Open the render_script document in the species folder and run that script (see README.md for additional instructions). It will automatically write the files to the correct folders in the repository. Remember not to 'run all' in the RMarkdown. You will get an error.

# 12. Check diagnostics

## The final section of the output doc will be model diagnostics. Check the output word doc (it will be in the 'Knitted_Markdowns' folder) for any issues and re-run the render_script if necessary after selecting different options in the RMarkdown (e.g., does the fitness variable need to be transformed? Is there overdispersion in the dispersal data?)

## You should also check the word doc for warning messages for the models (e.g., singularity in mixed models). These (if there are any) will print in the knitted doc next to the model results.

## If problems persist, the way the analysis is done may have to be changed (check in with Celina).

## Note that the current default description of the diagnostics is: "There was no evidence for lack of fit for either the fitness or the dispersal model." Change this as needed.


#############

# YOU ARE DONE! NO NEED TO FILL IN OR CHANGE ANYTHING BELOW.

############

```

## Statistical Analysis

```{r loaddata, echo = F}

# Set reading and writing paths; Do not change these paths
reading.path <- ("./")

writing.path <- ("../outputs/")

# load data file
d_phase1 <- read.csv(paste("Phase1_fitness_", Genus, ".", species, ".csv", sep = ""))
d_phase2 <- read.csv(paste("Phase2_dispersal_", Genus, ".", species, ".csv", sep = ""))

d_phase1 <- within(d_phase1, Temp.treatment <- relevel(as.factor(Temp.treatment), ref = "opt"))
d_phase2 <- within(d_phase2, Temp.treatment <- relevel(as.factor(Temp.treatment), ref = "opt"))

d_phase2 <- d_phase2 %>%
  { if (filter.dead) filter(., Survived == "Y") else . }

#modify
d_phase1 <- d_phase1 %>% filter(Temp.treatment != "lab")

d_phase2 <- d_phase2 %>% filter(Temp.rearing == "opt")
d_phase2 <- d_phase2 %>% filter(Temp.treatment!= "lab")
#sex filter
d_phase2 <- d_phase2 %>% filter(Sex == "M")


```

```{r writedata, echo = F}

# specify fitness and dispersal file names
fitpath <- paste("Phase1_fitness_", Genus, ".", species, ".csv", sep = "")
disppath <- paste("Phase2_dispersal_", Genus, ".", species, ".csv", sep = "")

data_path <- "../raw_data/" #path to save the data is to the raw_data folder

#save copy of final dfs used in this file to the raw_data folder
write.csv(d_phase1, paste(data_path, fitpath, sep = ""), row.names = FALSE)
write.csv(d_phase2, paste(data_path, disppath, sep = ""), row.names = FALSE)

```


### Does fitness change with temperature (phase 1)?

`r ifelse(fitness.transformed.normal == TRUE, print(paste(fitness.proxy, " was not normally distributed so we ", transform.description, "-transformed it. The transformed variable followed a normal distribution.", sep = "")), ifelse(fitness.normal == TRUE, print(paste(fitness.proxy, "was normally distributed.")), print(paste("The fitness proxy,", tolower(fitness.proxy), ", was a binary variable."))))`

```{r fitnessnormal, eval = fitness.normal, echo = F, fig.cap = paste("Figure 1. Histogram of fitness (", tolower(fitness.proxy), ") values.")}

# histogram showing fitness proxy was normally distributed

with(d_phase1, hist(Fitness, xlab = fitness.proxy, main = NULL))

```


```{r transformedfitnessnormal, eval = fitness.transformed.normal, echo = F, fig.cap = paste("Figure 1. Histogram of", transform.description, "-transformed", tolower(fitness.proxy), "values.")}

# histogram showing transformed fitness proxy was normally distributed

with(d_phase1, hist(fitness.transformation.function(Fitness), xlab = paste(transform.description, "-transformed ", tolower(fitness.proxy), sep = ""), main = NULL))

```

```{r fitnessplotnormal, eval = fitness.normal | fitness.transformed.normal, echo = F, message = F, fig.cap = paste("Figure 2. Boxplot of fitness (", tolower(fitness.proxy), "), as a function of temperature treatment. low = ", low.temp, "C, opt = ", opt.temp, "C, high = ", high.temp, "C.")}

# plot fitness data

ggplot(d_phase1, aes(x = Temp.treatment, y = Fitness, col = Temp.treatment))+
  geom_boxplot()+
  scale_x_discrete("Temperature treatment", limits = c("low", "opt", "high"))+
  scale_color_manual(values = c("darkgreen", "red", "blue"), guide = "none")+
  theme_classic(base_size = 15)

```

`r ifelse(fitness.binary == T, print(paste("To estimate the effects of temperature (a factor) on fitness we used a generalized linear mixed model with a logit link and", tolower(fitness.proxy), "as the binary response variable. We used the formula:")), paste("To estimate the effects of temperature (a factor) on fitness, we used a linear mixed model with", tolower(fitness.proxy), "as the response variable. We used the formula:"))`  

`r (print(fitness.formula))`

```{r fitnessmodelnormal, eval = fitness.normal | fitness.transformed.normal, echo = F}

# model testing effects of temperature on fitness

m_fitness <- lm(fitness.formula, data = d_phase1)

# output results for overall model and pairwise comparisons

output_fitness_Tempeffect <- anova(m_fitness)
output_fitness_summary <- summary(m_fitness)
output_fitness_emmeans <- emmeans(m_fitness, specs = pairwise ~ Temp.treatment, regrid = "response", adjust = "none")
output_fitness_contrasts <- output_fitness_emmeans$contrasts %>%
     summary(infer = TRUE)
```


```{r fitnesstablebinary, eval = fitness.binary, echo = F}

fittable <- with(d_phase1, table(Temp.treatment, Fitness))
knitr::kable(fittable, col.names = c(Fit0, Fit1),
 caption = paste("Table 1. Relationship between", tolower(fitness.proxy), "and temperature treatment.")) 


```


```{r fitnessmodelbinary, eval = fitness.binary, echo = F}

# model testing effects of temperature on fitness

m_fitness <- glmer(fitness.formula, family = binomial, data = d_phase1)

# output results for overall model and pairwise comparisons

output_fitness_Tempeffect <- drop1(m_fitness, test = "Chisq")
output_fitness_summary <- summary(m_fitness)
output_fitness_emmeans <- emmeans(m_fitness, specs = pairwise ~ Temp.treatment, type = "response", adjust = "none")
output_fitness_contrasts <- output_fitness_emmeans$contrasts %>%
     summary(infer = TRUE)

```

#### Results for main effect of temperature on fitness
```{r fitnessresultsmaineffect, echo = F}

output_fitness_Tempeffect

```

#### Post-hoc pairwise contrasts: differences in fitness between temperature treatments
```{r fitnessresultscontrasts, echo = F}

output_fitness_contrasts

```


### Does dispersal change with temperature (phase 2)?

`r ifelse(!overdispersion, print("To estimate the effect of temperature treatment (a factor) on dispersal, we used a generalized linear mixed model with a binomial error structure and a logit link. Dispersal (the number of dispersers vs non-dispersers) was the response variable. We used the formula:"), print("There was evidence of overdispersion in our dispersal data. Therefore, to estimate the effect of temperature treatment (a factor) on dispersal, we used a generalized linear mixed model with a negative binomial error structure. Dispersal (the number of dispersers vs non-dispersers) was the response variable. We used the formula:"))`  

`r (print(dispersal.formula))`


```{r dispersalmodelbinomial, eval = !overdispersion, echo = F}

# binomial model testing effects of temperature on dispersal

m_dispersal <- glmer(dispersal.formula, family = "binomial", data = d_phase2)

# output results for overall model and pairwise comparisons
output_dispersal_Tempeffect <- drop1(m_dispersal, test = "Chisq")
output_dispersal_summary <- summary(m_dispersal)
output_dispersal_emmeans <- emmeans(m_dispersal, specs = pairwise ~ Temp.treatment, type = "response", adjust = "none")
output_dispersal_contrasts <- output_dispersal_emmeans$contrasts %>%
  summary(infer = TRUE)


```

```{r dispersalmodelnegbinomial, eval = overdispersion, echo = F}

# negative binomial model testing effects of temperature on dispersal

m_dispersal <- glmer.nb(dispersal.formula, data = d_phase2)

# output results for overall model and pairwise comparisons
output_dispersal_Tempeffect <- drop1(m_dispersal, test = "Chisq")
output_dispersal_summary <- summary(m_dispersal)
output_dispersal_emmeans <- emmeans(m_dispersal, specs = pairwise ~ Temp.treatment, type = "response", adjust = "none")
output_dispersal_contrasts <- output_dispersal_emmeans$contrasts %>%
  summary(infer = TRUE)

```

```{r dispersalplot, echo = F, warning = F, fig.cap = paste("Figure 1. Dispersal as a function of temperature treatment. low = ", low.temp, "C, opt = ", opt.temp, "C, high = ", high.temp, "C. Error bars represent 95% confidence intervals predicted from GLMM.")}

# plot dispersal data

ggplot(as.data.frame(output_dispersal_emmeans$emmeans), aes(x = Temp.treatment, y = prob, col = Temp.treatment))+
  geom_point(size = 4)+
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0)+
  scale_x_discrete("Temperature treatment", limits = c("low", "opt", "high"))+
  scale_y_continuous("Dispersal probability")+
  scale_color_manual(values = c("darkgreen", "red", "blue"), guide = "none")+
  theme_classic(base_size = 15)

```

#### Results for main effect of temperature on dispersal
```{r dispersalresultsmaineffect, echo = F}

output_dispersal_Tempeffect

```

#### Post-hoc pairwise contrasts: differences in dispersal between temperature treatments
```{r dispersalresultscontrasts, echo = F}

output_dispersal_contrasts

```

### Model Diagnostics

There was no evidence for lack of fit for either the fitness or the dispersal model.

#### Diagnostic plots for fitness data

```{r modeldiagnosticsfitness, echo = F, fig.width = 8, fig.height = 7}

diagnostic_fitness <- simulateResiduals(fittedModel = m_fitness)
plot(diagnostic_fitness)
```

#### Diagnostic plots for dispersal data

```{r modeldiagnosticsdispersal, echo = F, fig.width = 8, fig.height = 7}

diagnostic_dispersal <- simulateResiduals(fittedModel = m_dispersal)
plot(diagnostic_dispersal)

```

```{r moderators, echo = F}

d_moderator <- data.frame(
  "Species" = paste(Genus, species),
  "Taxonomic group" = taxonomic.group,
  "Rearing.environment" = rearing.environment,
  "Competition.strength" = comp.strength,
  "Source" = source,
  "Habitat" = habitat,
  "Temperature.variation" = Tvariation,
  "Species mean body size" = species.body.size,
  "Passive.dispersal" = passive.dispersal,
  "Inbred.or.Mixed" = inbred,
  "Dispersal.assay.duration" = disp.assay.duration,
  "Fitness.proxy.type" = fitness.proxy.type
)

```

```{r writingoutputfiles, echo = F}

# fitness output

write.table(output_fitness_Tempeffect, paste(writing.path, "output_fitness_TempMain_", Genus, ".", species, ".txt", sep = ""))

write.table(output_fitness_summary$coefficients, paste(writing.path, "output_fitness_Summary_", Genus, ".", species, ".txt", sep = ""))

write.table(output_fitness_emmeans$emmeans, paste(writing.path, "output_fitness_emmeans_", Genus, ".", species, ".txt", sep = ""))

write.table(output_fitness_emmeans$contrasts, paste(writing.path, "output_fitness_contrasts_", Genus, ".", species, ".txt", sep = ""))


# dispersal output

write.table(output_dispersal_Tempeffect, paste(writing.path, "output_dispersal_TempMain_", Genus, ".", species, ".txt", sep = ""))

write.table(output_dispersal_summary$coefficients, paste(writing.path, "output_dispersal_Summary_", Genus, ".", species, ".txt", sep = ""))

write.table(output_dispersal_emmeans$emmeans, paste(writing.path, "output_dispersal_emmeans_", Genus, ".", species, ".txt", sep = ""))

write.table(output_dispersal_emmeans$contrasts, paste(writing.path, "output_dispersal_contrasts_", Genus, ".", species, ".txt", sep = ""))


# moderator output

write.table(d_moderator, paste(writing.path, "moderators_", Genus, ".", species, ".txt", sep = ""))

```
