---
title: "Temperature dependent dispersal: A distributed experiment"
author: "Peter Kamal & Emanuel A. Fronhofer"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r instructions, echo = F, warning = F}

# INSTRUCTIONS FOR RESEARCHERS

# Here, you'll find instructions and information to fill out about your species. Once you have finished following the instructions, go to the render_script in your species folder and run the code - this will automatically knit a word document to the shared Knitted_Markdowns folder. You don't need to make any changes to the analysis outside of this code chunk, but feel free to do so if it's appropriate for your species. If you make substantive changes to either of the models, please flag it for Celina.

# Note: Do not try to "run all" within R, you will get an error. Knitting the document from the RStudio menu will only save it to your species folder, not add it to the shared folder, so please use render_script.

# 0. Add names of researchers who worked on this species to author list above

# 1. Before starting, make sure your data files are named:
# Phase1_fitness_Genus.species.csv
# Phase2_dispersal_Genus.species.csv

## code below assumes .csv format (only in the code chunk named 'loaddata'; You can change this if you would rather not use .csv)


# 2. Install any missing libraries from the list below:
library(tidyverse)
library(lme4)
library(ggplot2)
library(emmeans)
library(DHARMa)

# 3. Fill out information about your species below

Genus = "Tetrahymena"
species = "thermophila_ISEM"
fitness.proxy = "r"
low.temp = 20 #value of low temperature treatment in Celsius
opt.temp = 36 #value of optimal temperature treatment in Celsius
high.temp = 38 #value of high temperature treatment in Celsius


# 4. Check that the data files have the following headers (order of columns is not important):

## phase1 : Temp.treatment (levels = low, opt, high), Temp.value (in degrees C), random effects (name as appropriate), Fitness
### It's important to give the column of the fitness variable the header "Fitness". You have already provided a description of the fitness proxy which will be used in the knitted document.

## phase 2:

### If your phase 2 data are formatted with mesocosms as rows and number of dispersers/number of non-dispersers as columns, then these headers:
#### Temp.treatment (levels = low, opt, high), Temp.value (in degrees C), random effects (name as appropriate), Num.disp, Num.nondisp (or alternatively, Num.disp and total N)

### If your phase 2 data are formatted with individuals as rows and a dispersed column (0s and 1s) giving dispersal status for each individual, then these headers:
#### Temp.treatment (levels = low, opt, high), Temp.value (in degrees C), random effects (name as appropriate), Dispersed (0/1)

# note that the datasets can also include fixed effects and phenotypes, as appropriate for your study


# 5. Type of fitness data

## Answer questions below about the fitness proxy

### Is your fitness variable binary? (choose one)
fitness.binary = FALSE

### Is the fitness variable continuous normal (without transformation)? (choose one)
fitness.normal = TRUE

### Is the fitness variable normal after transformation? (choose one; choose NA if you said "TRUE" to either question above)
fitness.transformed.normal = NA

## If fitness variable is continuous normal AFTER transformation, choose transformation used (just write it in if not on this list):

transform.description <- "square-root/log/log10" # leave as-is if not applicable
fitness.transformation.function <- function(x) {return (x)} # give transformation formula here; replace XYZ with transformation e.g., sqrt(x) or log(x + 1), etc., where x will be Fitness, below

## note: if fitness (or transformed fitness) does not have a binomial or normal distribution, check in with Celina.


# 6. Structure of dispersal data

## Is there overdispersion in the dispersal data? If you do not yet know, choose FALSE and this will be dealt with in step 12.

overdispersion = FALSE


# 7. Dealing with mortality in dispersal assay

## note on mortality: everyone will deal with mortality during the dispersal assay in a slightly different way. If individuals are rows in your dataset, you should have a row that indicates "Survived" (Y/N) through the entire experiment. In this case, set filter.dead = TRUE below. If mesocosms are rows in your dataset, you may have a column indicating the number of individuals that died. In this case, dead individuals can be excluded directly in the dispersal formula (e.g., (Num.disp, TotalN-(Num.disp+Num.dead))).

## Make sure to describe how you dealt with mortality in the methods below.

## If you are unsure how to handle mortality in your dataset, check with Celina.

## Do rows represent individuals in the dataset and are there dead individuals that should be removed from the dispersal data?

filter.dead = FALSE # choose NA if this does not apply to your species


# 8. Model formulae

## Formula for fitness: (fill in as appropriate)

fitness.formula <- "Fitness ~ Temp.treatment + (1|Replicate)"

### notes:
#### TRANSFORM FITNESS VARIABLE (if needed) *WITHIN* THE FORMULA; doing it within the formula is important for contrasts below  
#### other.fixed.effects does not apply to every species; include as appropriate; fixed effects should be standardized
#### include as many random effects as appropriate
#### Fitness can be normal or binary, including 0s and 1s or cbind(number.survived, number.dead)

### If fitness is binary: (ignore if fitness is not binary)

### What do Fitness = 0 and Fitness = 1 mean?

Fit0 <- "Died/Survived" # or fill in other
Fit1 <- "Died/Survived" # or fill in other

## Formula for dispersal: (fill in as appropriate)
  
dispersal.formula <- "cbind(Num.disp, Num.nondisp) ~ Temp.treatment + (1|Replicate)"

### notes
#### other.fixed.effects does not apply to every species; include as appropriate; fixed effects should be standardized
#### include as many random effects as appropriate
#### Num.disp/Num.nondisp can be replaced with column "Dispersed" with values 0 and 1, if individuals are rows in your dataset
#### remember to deal with mortality directly in dispersl formula if appropriate


# 9. Moderators

## We are collecting data on all proposed moderators; we will use a subset of these in the analysis
## choose option on right hand side
## see description of moderators and levels in README.md in project repo

taxonomic.group <- "ciliate"
rearing.environment <- "commonT"
comp.strength <- "moderate"
source <- "lab" # lab population or field caught
habitat <- "aquatic"
Tvariation <- "homogenous"
species.body.size <- 0.05 # mean body length in mm
passive.dispersal <- "No"
inbred <- "inbred line"
disp.assay.duration <- 1 # order of magnitude for dispersal assay duration / generation time, i.e., approximate number of generations the dispersal assay lasted
fitness.proxy.type <- "r"

# 10. Fill out methods information in text below.

```

# Supplementary methods for *`r (print(paste(Genus, species)))`*

## Study organism

* **Name and strain of species:** *Tetrahymena thermophila* (strain SB3539)
* **Basic life history:** Aquatic protist, active dispersal via swimming, generation time of ~ 4 hours
* **Temperature range:** Lives originally in streams, ponds and lakes on the Atlantic coast of North America. Our strain has been kept in stock for at least six years (around 13,000 generations) at 20$^\circ$C.
* **Source:** Cornell Tetrahymena Stock Center
* **Holding conditions before experiment**: For each experiment, we used individuals from our running batch cultures, kept at 20$^\circ$C in a dark incubator in sterile medium made up of mineral water, ground lettuce, and the bacterial species *Serratia fonticola*. The medium was filtered to provide homogeneous spatial structure within the vessel. Whenever we add medium to a batch culture or a test tube, that medium is enriched with a three-day old culture of the bacterium *Serratia fonticola* at a ratio of 9:1 (medium:bacteria). No further selection for the experimental individuals was conducted. 

## Study design

### Temperature treatments

State the temperatures used for Topt, Thot, Tcold, and explain how they were chosen: We used Topt = 36C, Thot = 38C, and Tcold = 20C. Tcold is the culture temperature, Topt is the temperature at which the growth rate is maximal, Thot is the temperature at which some growth is still possible, but it is on the descending flank of the TPC. There was no daily variation in temperature. 

### Phase 1: Fitness assay

* **Fitness proxy:** Maximum growth rate per replicate and temperature, estimated as suggested with the easy linear fit in the growth rates package based on time series data at different temperatures starting from low densities (~100 individuals per ml).
* **Experimental setup:** Experiments were conducted in plastic vials (Sarstedt), filled with 20ml of the bacterized medium described above and kept in dark incubators.
* **Replication:** 5 experimental replicates per temperature. To avoid pseudoreplication, we switched incubators randomly every two days.
* **Food provided:** Every two days, we replenished 10% of the culture with fresh medium that was enriched with bacteria at the same 9:1 ratio.
* **Duration**: The experiment lasted 9 days (until the populations at the coldest temperature had reached carrying capacity).
* **Phenotypic assays:** We measured densities and phenotypes by recording videos under a stereoscope and analyzing those videos with the `bemovi` software. That yielded (apart from densities) measurements of cell size as well as swimming speed.

### Phase 2: Dispersal assay

* **Setup: ** We conducted the experiment in the same plastic vials filled with 20ml of total culture medium, but now as a 2-patch system connected by a 7cm long silicon tube that we could close with plastic clamps. We consider this silicone tube to be hostile due to the lack of oxygen (which *Tetrahymena* are very sensitive to). 
* **Dispersal: ** We measured emigration rates by opening the connections between origin and target patch, closing them again after 3 hours, and sampling in both patches. We estimated dispersal rate as the ratio of individuals in the target patch to the total number of individuals in both patches combined.
* **Replication: ** We replicated each dispersal trial at each temperature 5 times (yielding 15 total experimental units). Individuals were not the same as the ones used in the fitness assay, but came from the same clonal population as above. All five replicates of each temperature were assayed at the same time in the same incubator.
* **Inoculation: ** We used the fitness assay to determine which inoculation density would produce the highest variation in estimated fitness across temperatures (see below). We chose a density of 1,500 individuals per ml, which is between 25 and 50% of K for all temperatures. In the morning of the experiment, we measured our stock and determined the necessary amount to add to a 20ml tube to reach this density. Accordingly, we inoculated the origin patch with 5ml of *Tetrahymena* stock and 15ml of bacterized medium (see above), and the target patch with 20ml of bacterized medium.
* **Timing: ** We estimate *Tetrahymena's* generation time to be roughly 4 hours. Therefore, we let the cultures acclimate with the gates closed in the incubator for 45 minutes, and then opened the gates for 3 hours, after which we closed them and sampled.
* **Sampling and phenotyping: ** As before, we sampled out of the cultures and recorded videos under a stereoscope, which gave us both density and phenotypic estimates. To reduce sampling error given the possible low densities in the target patch, we used the mean across three technical replicates for each sample.
* **Mortality: ** was not recorded.



```{r instructions2, echo = F}

# 11. Knit RMarkdown

## Now it is time to knit the RMarkdown document. Open the render_script document in the species folder and run that script (see README.md for additional instructions). It will automatically write the files to the correct folders in the repository. Remember not to 'run all' in the RMarkdown. You will get an error.

# 12. Check diagnostics

## The final section of the output doc will be model diagnostics. Check the output word doc (it will be in the 'Knitted_Markdowns' folder) for any issues and re-run the render_script if necessary after selecting different options in the RMarkdown (e.g., does the fitness variable need to be transformed? Is there overdispersion in the dispersal data?)

## You should also check the word doc for warning messages for the models (e.g., singularity in mixed models). These (if there are any) will print in the knitted doc next to the model results.

## If problems persist, the way the analysis is done may have to be changed (check in with Celina).

## Note that the current default description of the diagnostics is: "There was no evidence for lack of fit for either the fitness or the dispersal model." Change this as needed.


#############

# YOU ARE DONE! NO NEED TO FILL IN OR CHANGE ANYTHING BELOW.

############

```

## Statistical Analysis

```{r loaddata, echo = F}

# Set reading and writing paths; Do not change these paths
reading.path <- ("./")

writing.path <- ("../outputs/")

# load data file
d_phase1 <- read.csv(paste("Phase1_fitness_", Genus, ".", species, ".csv", sep = ""))
d_phase2 <- read.csv(paste("Phase2_dispersal_", Genus, ".", species, ".csv", sep = ""))

d_phase1 <- within(d_phase1, Temp.treatment <- relevel(as.factor(Temp.treatment), ref = "opt"))
d_phase2 <- within(d_phase2, Temp.treatment <- relevel(as.factor(Temp.treatment), ref = "opt"))

d_phase2 <- d_phase2 %>%
  { if (filter.dead) filter(., Survived == "Y") else . }

```

```{r writedata, echo = F}

# specify fitness and dispersal file names
fitpath <- paste("Phase1_fitness_", Genus, ".", species, ".csv", sep = "")
disppath <- paste("Phase2_dispersal_", Genus, ".", species, ".csv", sep = "")

data_path <- "../raw_data/" #path to save the data is to the raw_data folder

#save copy of final dfs used in this file to the raw_data folder
write.csv(d_phase1, paste(data_path, fitpath, sep = ""), row.names = FALSE)
write.csv(d_phase2, paste(data_path, disppath, sep = ""), row.names = FALSE)

```


### Does fitness change with temperature (phase 1)?

`r ifelse(fitness.transformed.normal == TRUE, print(paste(fitness.proxy, " was not normally distributed so we ", transform.description, "-transformed it. The transformed variable followed a normal distribution.", sep = "")), ifelse(fitness.normal == TRUE, print(paste(fitness.proxy, "was normally distributed.")), print(paste("The fitness proxy,", tolower(fitness.proxy), ", was a binary variable."))))`

```{r fitnessnormal, eval = fitness.normal, echo = F, fig.cap = paste("Figure 1. Histogram of fitness (", tolower(fitness.proxy), ") values.")}

# histogram showing fitness proxy was normally distributed

with(d_phase1, hist(Fitness, xlab = fitness.proxy, main = NULL))

```


```{r transformedfitnessnormal, eval = fitness.transformed.normal, echo = F, fig.cap = paste("Figure 1. Histogram of", transform.description, "-transformed", tolower(fitness.proxy), "values.")}

# histogram showing transformed fitness proxy was normally distributed

with(d_phase1, hist(fitness.transformation.function(Fitness), xlab = paste(transform.description, "-transformed ", tolower(fitness.proxy), sep = ""), main = NULL))

```

```{r fitnessplotnormal, eval = fitness.normal | fitness.transformed.normal, echo = F, message = F, fig.cap = paste("Figure 2. Boxplot of fitness (", tolower(fitness.proxy), "), as a function of temperature treatment. low = ", low.temp, "C, opt = ", opt.temp, "C, high = ", high.temp, "C.")}

# plot fitness data

ggplot(d_phase1, aes(x = Temp.treatment, y = Fitness, col = Temp.treatment))+
  geom_boxplot()+
  scale_x_discrete("Temperature treatment", limits = c("low", "opt", "high"))+
  scale_color_manual(values = c("darkgreen", "red", "blue"), guide = "none")+
  theme_classic(base_size = 15)

```

`r ifelse(fitness.binary == T, print(paste("To estimate the effects of temperature (a factor) on fitness we used a generalized linear mixed model with a logit link and", tolower(fitness.proxy), "as the binary response variable. We used the formula:")), paste("To estimate the effects of temperature (a factor) on fitness, we used a linear mixed model with", tolower(fitness.proxy), "as the response variable. We used the formula:"))`  

`r (print(fitness.formula))`

```{r fitnessmodelnormal, eval = fitness.normal | fitness.transformed.normal, echo = F}

# model testing effects of temperature on fitness

m_fitness <- lmer(fitness.formula, data = d_phase1)

# output results for overall model and pairwise comparisons

output_fitness_Tempeffect <- anova(m_fitness)
output_fitness_summary <- summary(m_fitness)
output_fitness_emmeans <- emmeans(m_fitness, specs = pairwise ~ Temp.treatment, regrid = "response", adjust = "none")
output_fitness_contrasts <- output_fitness_emmeans$contrasts %>%
     summary(infer = TRUE)
```


```{r fitnesstablebinary, eval = fitness.binary, echo = F}

fittable <- with(d_phase1, table(Temp.treatment, Fitness))
knitr::kable(fittable, col.names = c(Fit0, Fit1),
 caption = paste("Table 1. Relationship between", tolower(fitness.proxy), "and temperature treatment.")) 


```


```{r fitnessmodelbinary, eval = fitness.binary, echo = F}

# model testing effects of temperature on fitness

m_fitness <- glmer(fitness.formula, family = binomial, data = d_phase1)

# output results for overall model and pairwise comparisons

output_fitness_Tempeffect <- drop1(m_fitness, test = "Chisq")
output_fitness_summary <- summary(m_fitness)
output_fitness_emmeans <- emmeans(m_fitness, specs = pairwise ~ Temp.treatment, type = "response", adjust = "none")
output_fitness_contrasts <- output_fitness_emmeans$contrasts %>%
     summary(infer = TRUE)

```

#### Results for main effect of temperature on fitness
```{r fitnessresultsmaineffect, echo = F}

output_fitness_Tempeffect

```

#### Post-hoc pairwise contrasts: differences in fitness between temperature treatments
```{r fitnessresultscontrasts, echo = F}

output_fitness_contrasts

```


### Does dispersal change with temperature (phase 2)?

`r ifelse(!overdispersion, print("To estimate the effect of temperature treatment (a factor) on dispersal, we used a generalized linear mixed model with a binomial error structure and a logit link. Dispersal (the number of dispersers vs non-dispersers) was the response variable. We used the formula:"), print("There was evidence of overdispersion in our dispersal data. Therefore, to estimate the effect of temperature treatment (a factor) on dispersal, we used a generalized linear mixed model with a negative binomial error structure. Dispersal (the number of dispersers vs non-dispersers) was the response variable. We used the formula:"))`  

`r (print(dispersal.formula))`


```{r dispersalmodelbinomial, eval = !overdispersion, echo = F}

# binomial model testing effects of temperature on dispersal

m_dispersal <- glmer(dispersal.formula, family = "binomial", data = d_phase2)

# output results for overall model and pairwise comparisons
output_dispersal_Tempeffect <- drop1(m_dispersal, test = "Chisq")
output_dispersal_summary <- summary(m_dispersal)
output_dispersal_emmeans <- emmeans(m_dispersal, specs = pairwise ~ Temp.treatment, type = "response", adjust = "none")
output_dispersal_contrasts <- output_dispersal_emmeans$contrasts %>%
  summary(infer = TRUE)


```

```{r dispersalmodelnegbinomial, eval = overdispersion, echo = F}

# negative binomial model testing effects of temperature on dispersal

m_dispersal <- glmer.nb(dispersal.formula, data = d_phase2)

# output results for overall model and pairwise comparisons
output_dispersal_Tempeffect <- drop1(m_dispersal, test = "Chisq")
output_dispersal_summary <- summary(m_dispersal)
output_dispersal_emmeans <- emmeans(m_dispersal, specs = pairwise ~ Temp.treatment, type = "response", adjust = "none")
output_dispersal_contrasts <- output_dispersal_emmeans$contrasts %>%
  summary(infer = TRUE)

```

```{r dispersalplot, echo = F, warning = F, fig.cap = paste("Figure 1. Dispersal as a function of temperature treatment. low = ", low.temp, "C, opt = ", opt.temp, "C, high = ", high.temp, "C. Error bars represent 95% confidence intervals predicted from GLMM.")}

# plot dispersal data

ggplot(as.data.frame(output_dispersal_emmeans$emmeans), aes(x = Temp.treatment, y = prob, col = Temp.treatment))+
  geom_point(size = 4)+
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0)+
  scale_x_discrete("Temperature treatment", limits = c("low", "opt", "high"))+
  scale_y_continuous("Dispersal probability")+
  scale_color_manual(values = c("darkgreen", "red", "blue"), guide = "none")+
  theme_classic(base_size = 15)

```

#### Results for main effect of temperature on dispersal
```{r dispersalresultsmaineffect, echo = F}

output_dispersal_Tempeffect

```

#### Post-hoc pairwise contrasts: differences in dispersal between temperature treatments
```{r dispersalresultscontrasts, echo = F}

output_dispersal_contrasts

```

### Model Diagnostics

There was no evidence for lack of fit for either the fitness or the dispersal model.

#### Diagnostic plots for fitness data

```{r modeldiagnosticsfitness, echo = F, fig.width = 8, fig.height = 7}

diagnostic_fitness <- simulateResiduals(fittedModel = m_fitness)
plot(diagnostic_fitness)
```

#### Diagnostic plots for dispersal data

```{r modeldiagnosticsdispersal, echo = F, fig.width = 8, fig.height = 7}

diagnostic_dispersal <- simulateResiduals(fittedModel = m_dispersal)
plot(diagnostic_dispersal)
```

```{r moderators, echo = F}

d_moderator <- data.frame(
  "Species" = paste(Genus, species),
  "Taxonomic group" = taxonomic.group,
  "Rearing.environment" = rearing.environment,
  "Competition.strength" = comp.strength,
  "Source" = source,
  "Habitat" = habitat,
  "Temperature.variation" = Tvariation,
  "Species mean body size" = species.body.size,
  "Passive.dispersal" = passive.dispersal,
  "Inbred.or.Mixed" = inbred,
  "Dispersal.assay.duration" = disp.assay.duration,
  "Fitness.proxy.type" = fitness.proxy.type
)

```

```{r writingoutputfiles, echo = F}

# fitness output

write.table(output_fitness_Tempeffect, paste(writing.path, "output_fitness_TempMain_", Genus, ".", species, ".txt", sep = ""))

write.table(output_fitness_summary$coefficients, paste(writing.path, "output_fitness_Summary_", Genus, ".", species, ".txt", sep = ""))

write.table(output_fitness_emmeans$emmeans, paste(writing.path, "output_fitness_emmeans_", Genus, ".", species, ".txt", sep = ""))

write.table(output_fitness_emmeans$contrasts, paste(writing.path, "output_fitness_contrasts_", Genus, ".", species, ".txt", sep = ""))


# dispersal output

write.table(output_dispersal_Tempeffect, paste(writing.path, "output_dispersal_TempMain_", Genus, ".", species, ".txt", sep = ""))

write.table(output_dispersal_summary$coefficients, paste(writing.path, "output_dispersal_Summary_", Genus, ".", species, ".txt", sep = ""))

write.table(output_dispersal_emmeans$emmeans, paste(writing.path, "output_dispersal_emmeans_", Genus, ".", species, ".txt", sep = ""))

write.table(output_dispersal_emmeans$contrasts, paste(writing.path, "output_dispersal_contrasts_", Genus, ".", species, ".txt", sep = ""))


# moderator output

write.table(d_moderator, paste(writing.path, "moderators_", Genus, ".", species, ".txt", sep = ""))

```
